---
title: "Stats170_Hwk_1"
author: "Shiqi Liang, Ingrid Wijaya, Jessica Wong"
output: pdf_document
---
### GROUP MEMBERS
Shiqi Liang \
Ingrid Wijaya \
Jessica Wong 

### DATA SOURCE

FRED Total Business Sales(TOTBUSSMNSA) \
FRED Average Hourly Earnings of Production and Nonsupervisory Employees, Total Private (CEU0500000008) \
FRED Unemployment Rate (UNRATE) 

### Loading packages
```{r}
library(ggplot2)
library(Quandl)
```

### Loaing data from Quandl
Since we are only considering none-Covid years, we have decided to only look at data before Jan 2020.
```{r}
biz_sales = Quandl(code="FRED/TOTBUSSMNSA",
            type="ts",
            collapse="monthly",
            order="asc",
            end_date="2019-12-31",
            meta=TRUE)

avg_hr_sal = Quandl(code="FRED/CEU0500000008",
            type="ts",
            collapse="monthly",
            order="asc",
            end_date="2019-12-31",
            meta=TRUE)

unrate =  Quandl(code="FRED/UNRATE",
            type="ts",
            collapse="monthly",
            order="asc",
            end_date="2019-12-31",
            meta=TRUE)

```

### Creating a dataframe with the intersect, and create training and testing set
```{r}
full = ts.intersect(biz_sales,avg_hr_sal,unrate)
train = window(full,start = c(1992,1), end = c(2018,12))
test = window(full,start = c(2019,1), end=c(2019,12))
```

### Creating a table that stores relavent information
```{r}
info = data.frame(name = c('biz_sales','avg_hr_sal','unrate'),
                  variable_short_name =      c('FRED/TOTBUSSMNSA','FRED/CEU0500000008','FRED/UNRATE'),
                  description = c('Total business sales within the United States by month','average hourly earnings of production and nonsupervisory employees in the private sector of the United Stats by month', 'unemployment rate in the United States by month'),
                  training_time_interval = c('1992-1 to 2018-12','1992-1 to 2018-12','1992-1 to 2018-12'),
                  testing_time_interval = c('2019-1 to 2019-12','2019-1 to 2019-12','2019-1 to 2019-12'))

info
```

### Part D:
```{r}
ts.plot(train[,'biz_sales'],main = 'total business sales',ylab = 'sales in millions of dollars')

season <- cycle(train[,'biz_sales'])
season.factor <-factor(season)
ggplot()+
  geom_boxplot(mapping = aes(x = season.factor,y=train[,'biz_sales']))+labs(x = 'month',y = 'business sales in millions of dollars')
```
There is a strong upward trend. We can see  small seasonal component in business sales, although it is important to point out that in the seasonal box plots the boxes overlap each other, so we can not say the seasonal component is very obvious.

```{r}
ts.plot(train[,'avg_hr_sal'],main = 'Average Hourly Wage',ylab = 'Dollars per Hour')

season <- cycle(train[,'avg_hr_sal'])
season.factor <-factor(season)
ggplot()+
  geom_boxplot(mapping = aes(x = season.factor,y=train[,'avg_hr_sal']))+labs(x = 'month',y = 'Average Hourly Wage')
```
There is a strong upward trend and no obvious seasonal component is present. This is quiet expected since hourly wages are expected to remain relatively stable through out the year. 

```{r}
ts.plot(train[,'unrate'],main = 'Unemployment Rate',ylab = 'Percent')

season <- cycle(train[,'unrate'])
season.factor <-factor(season)
ggplot()+
  geom_boxplot(mapping = aes(x = season.factor,y=train[,'unrate']))+labs(x = 'month',y = 'Unemployment Rate')
```
There seems to be a few spikes in unemployment rate (possibly caused by economic crisis) but we can observe that there isn't a seasonal component in unemployment rates, possibly due to the fact that employment rate should be relatively stable throughout the year (except in times of economic crisis)

### Part E 

1. Dependent variable 
  - Business Sales  

2. Independent variable  
  - Average Hourly Earnings of Production and Nonsupervisory Employees      
  -Unemployment Rate

Referring to the time plot of business sales, due to increasing trend and increasing seasonality, we chose to do multiplicative decomposition. 

```{r}
sales_decomp = decompose(train[,'biz_sales'], type = "mult") 
plot(sales_decomp)

sales_trend = sales_decomp$trend
sales_season = sales_decomp$seasonal
sales_ran = sales_decomp$random

plot(sales_trend,main="Trend Component of Business Sales",ylab="sales in millions of dollars")
plot(sales_season,main="Seasonal component of Business Sales",ylab="BS-trend")
plot(sales_ran,main="Random component of Business Sales", ylab="BS-trend-seasonal")
```

From the plots above, some notable points are:  

1. The trend plot shows that there is a dip in business sales between 2007 and 2010,
which can indicate several causes such as an economic crisis or recession .   

2. The random plot shows that it is randomly fluctuating around a constant, indicating that multiplicative decomposition is appropriate. We observed that between 2007 and 2010, the random component has a strikingly large fluctuation compared to the rest, the same time period business sales decreased rapidly.

### Part F
```{r}
## random part 
new_sales_ran = window(sales_ran, start=c(1992,7), end=c(2018,6))
acf(new_sales_ran, main="ACF of Random Component of Business Sales")
```

Referring to the ACF plot, there are several autocorrelation that are above the
blue band, indicating that they are statistically significant.
